### 2022-01-28 plos_latex_template.tex
Recent edits

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4},
Heather Turner\textsuperscript{5}, %\ddag},
Dorothea Hug Peter\textsuperscript{6}, %\ddag},
Matt Bannert\textsuperscript{7},
Batool Almazrouq\textsuperscript{8,9},
Liz Hare\textsuperscript{10},
Laura Ación\textsuperscript{11},
Juan Pablo Narváez-Gómez\textsuperscript{12},
Marcela Alfaro Córdoba\textsuperscript{13},
Federico Marini\textsuperscript{14},
Rita Giordano\textsuperscript{15},
Natalia Morandeira\textsuperscript{16},
Adithi Upadhya\textsuperscript{17},
Joselyn Chávez\textsuperscript{18},
Anicet Ebou\textsuperscript{19},
Silvia Canelón\textsuperscript{20},
Janani Ravi\textsuperscript{21}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{5} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{8} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{9} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{13} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{17} ILK Labs/Bengaluru, India - 560046
\\
\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{19} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and network within an academic or technical field, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces 10 simple rules to host an inclusive conference. 
The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been mainly written for potential organizers or selection committees of conferences, and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from rich countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. let's try not to forget this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: pillar rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. From left to right, "the pillars" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team". Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial". To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
}
\caption{Schematic diagram of the rules organized in three groups: the pillars (Rules 1, 2, and 3), design rules (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

@@

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of preferred pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
Open the schedule for events organized by thematic groups and communities (e.g. taskforce meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
Furthermore, tasks such as receiving and responding to code of conduct reports can be emotionally intense work and should be additionally rewarded.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

@@

The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges (e.g. `manels' and all-white panels \cite{else_how_2019}).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international meetings, the linguistic diversity of the participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpages, and communications. 
While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege \cite{amanoTenTipsOvercoming2021}. 
Non-native English speakers may miss opportunities to attend or to actively participate in conferences (e.g. presenting, asking questions, or partaking in discussions),
and conferences may in turn miss innovative contributions.

@@

Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups–in all communications \cite{hallDesigningDiversityInclusion2019}. 
Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
Do not expect minoritized people to teach you and accept feedback without being offended.
Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms.

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was the useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly. From left to right: Margot wearing scarves with the colors of MiR (Minorities in R https://mircommunity.com/), R-Ladies https://rladies.org/, AfricaR https://africa-r.org/, and the LGBTQIA+ and transgender communities. Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license} 
\label{fig:marmots}
\end{figure}
%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves. From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the virtual component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards). Depending on the country, services like OpenCollective (https://opencollective.com/) can be very useful to do these transfers.

@@

\begin{itemize}
    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization
    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item YBS: Conceptualization, Writing--major reviews and edits, Resources
    \item HT: Conceptualization, Writing--major reviews and edits, Resources
    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
    \item MB: Conceptualization
    \item BA: Writing--major reviews and edits, Resources, Sharing rules
    \item LH: Writing--major reviews and edits, Resources, Sharing rules
    \item LA: Writing--major reviews and edits, Resources
    \item JPN: Writing--major reviews and edits, Visualization
    \item MAC: Writing--major reviews and edits
    \item FM: Writing--major reviews and edits
    \item RG: Writing--major reviews and edits
    \item NM: Sharing rules
    \item AU: Sharing rules
    \item JC: Sharing rules
    \item AE: Sharing rules
    \item SC: Sharing rules
    \item JR: Conceptualization, Writing--major reviews and edits
\end{itemize}

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team/}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.

Heather Turner was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR! 2021.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5,6},
Heather Turner\textsuperscript{7}, %\ddag},
Dorothea Hug Peter\textsuperscript{8}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{9},
Matt Bannert\textsuperscript{10},
Batool Almazrouq\textsuperscript{11,12},
Liz Hare\textsuperscript{13},
Laura Ación\textsuperscript{5,14},
Juan Pablo Narváez-Gómez\textsuperscript{15},
Marcela Alfaro Córdoba\textsuperscript{16},
Federico Marini\textsuperscript{17},
Rita Giordano\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Anicet Ebou\textsuperscript{21},
Silvia Canelón\textsuperscript{22},
Janani Ravi\textsuperscript{23}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{19} ILK Labs/Bengaluru, Karnataka 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{22} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference. 
The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been mainly written for potential organizers or selection committees of conferences and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: pillar rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. From left to right, "the pillars" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team". Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial". To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
}
\caption{Schematic diagram of the rules organized in three groups: the pillars (Rules 1, 2, and 3), design rules (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

@@

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of preferred pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
Furthermore, tasks such as receiving and responding to CoC reports can be emotionally intense work and should be additionally rewarded.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

@@

The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international meetings, the linguistic diversity of the participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, websites, and communications. 
While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege \cite{amanoTenTipsOvercoming2021}. 
Non-native English speakers may miss opportunities to attend or to actively participate in conferences (e.g. presenting, asking questions, or partaking in discussions),
and conferences may in turn miss innovative contributions.

@@

Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}. 
Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
Do not expect minoritized people to teach you and accept feedback without being offended.
Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms.

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was the useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly. From left to right: Margot wearing scarves with the colors of MiR (Minorities in R \url{https://mircommunity.com}), R-Ladies \url{https://rladies.org}, AfricaR \url{https://africa-r.org}, and the LGBTQIA+ and transgender communities. Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license (\url{https://github.com/useRconf/visuals-2021}).}
\label{fig:marmots}
\end{figure}
%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves. From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the virtual component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards). Depending on the country, services like OpenCollective (\url{https://opencollective.com}) can be very useful to do these transfers.

@@

\begin{itemize}
    \item RJ: Project administration, Supervision, Writing original draft, Conceptualization, Writing--major reviews and edits, Resources, Visualization
    \item AST: Project administration, Supervision, Conceptualization,  Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item SM: Conceptualization, Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item YBS: Conceptualization, Writing--major reviews and edits, Resources
    \item HT: Conceptualization, Writing--major reviews and edits, Resources
    \item DHP: Conceptualization, Writing--major reviews and edits, Sharing rules
    \item NSM: Writing--major reviews and edits, Resources, Visualization, Sharing rules
    \item MB: Conceptualization
    \item BA: Writing--major reviews and edits, Resources, Sharing rules
    \item LH: Writing--major reviews and edits, Resources, Sharing rules
    \item LA: Writing--major reviews and edits, Resources
    \item JPN: Writing--major reviews and edits, Visualization
    \item MAC: Writing--major reviews and edits
    \item FM: Writing--major reviews and edits
    \item RG: Writing--major reviews and edits
    \item AU: Sharing rules
    \item JC: Sharing rules
    \item AE: Sharing rules
    \item SC: Sharing rules
    \item JR: Conceptualization, Writing--major reviews and edits
\end{itemize}

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.

\section*{Funding}
HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR! 2021.
YBS was supported by the Instituto Nacional de Tecnología Agropecuaria, Argentina during the organization of useR! 2021.
... ...
